---
title: "Covid19 vaccine study Fall2021"
author: "April Galvan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

The bar chart shows that the Males who took the placebo got sick more than the ones Males who took the Covid91. We can see that the pink bar for placebo is almost doubled compared to the drug. This indicates the drug is affective for the Males group.

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
Based on the numerical summaries, 33.11% of the Males who got sick had the treatment and 66.89% of the Males who got sick took the placebo. This is a 2 to 1 difference and lets us know that the treatment is in fact working.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
MalesMod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small (2.2e-16). This indicats that we need to reject the null hypothesis and that there is a difference between the Males who took the placebo and Males who took the drug.

```{r}
MaleEstimate <- MalesMod$estimate
MaleLow <- MalesMod$conf.int[1]
MaleHigh <- MalesMod$conf.int[2]
MaleProb <- MaleEstimate / (1+MaleEstimate)
MaleLowProb <- MaleLow / (1+MaleLow)
MaleLowEffect <- MaleLowProb * 100
MaleHighProb <- MaleHigh / (1+MaleHigh)
MaleHighEffect <- MaleHighProb * 100
MaleEffect <- MaleProb * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Males is `r MaleEffect`. The confidence interval for males is `r MaleLowEffect` to `r MaleHighEffect`.

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According to the barchart, this vaccine is not as effective for Females. The blue bar in the chart indicates that out of the Females who are sick more than 50% had gotten the vaccine. 

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

## Inferential  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
FemalesMod <- fisher.test(table1)
```

# Drug Users

# LGBTQ

# Conclusions and recommendations

