---
title: "Covid19 vaccine study Fall2021"
author: "April Galvan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

The bar chart shows that the Males who took the placebo got sick more than the ones Males who took the Covid91. We can see that the pink bar for placebo is almost doubled compared to the drug. This indicates the drug is effective for the Males group.

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
Based on the numerical summaries, 33.11% of the Males who got sick had the treatment and 66.89% of the Males who got sick took the placebo. This is a 2 to 1 difference and lets us know that the treatment is in fact working.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
MalesMod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small (2.2e-16). This indicates that we need to reject the null hypothesis and that there is a difference between the Males who took the placebo and Males who took the drug.

## Calculating Effectiveness
```{r}
MaleEstimate <- MalesMod$estimate
MaleLow <- MalesMod$conf.int[1]
MaleHigh <- MalesMod$conf.int[2]
MaleProb <- MaleEstimate / (1+MaleEstimate)
MaleLowProb <- MaleLow / (1+MaleLow)
MaleLowEffect <- MaleLowProb * 100
MaleHighProb <- MaleHigh / (1+MaleHigh)
MaleHighEffect <- MaleHighProb * 100
MaleEffect <- MaleProb * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Males is `r MaleEffect`. The confidence interval for males is `r MaleLowEffect` to `r MaleHighEffect`.

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According to the barchart, this vaccine is not as effective for Females. The blue bar in the chart indicates that out of the Females who are sick more than 50% had gotten the vaccine. 

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

This table indicates that the Females who got sick, 58% of them took the drug versus 42% of the sick Females that got the placebo. Based on this information, we can see that the drug seems to not be as effective for Females.

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
FemalesMod <- fisher.test(table1)
```

Since our p-value is very small, we have to reject the null hypothesis. This indicates that there is a difference between our data. In this case, it looks as if the data is different in a bad way, meaning the drug if not effective compared to the placebo.

## Calculating Effectiveness

```{r}
FemaleEstimate <- FemalesMod$estimate
FemaleLow <- FemalesMod$conf.int[1]
FemaleHigh <- FemalesMod$conf.int[2]
FemaleProb <- FemaleEstimate / (1+FemaleEstimate)
FemaleLowProb <- FemaleLow / (1+FemaleLow)
FemaleLowEffect <- FemaleLowProb * 100
FemaleHighProb <- FemaleHigh / (1+FemaleHigh)
FemaleHighEffect <- FemaleHighProb * 100
FemaleEffect <- FemaleProb * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Females is `r FemaleEffect`. The confidence interval for Females is `r FemaleLowEffect` to `r FemaleHighEffect`.

# Drug Users

```{r}
DrugUser <- filter(Covid91, DrugUser == "yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

The bar chart shows that the Drug Users who took the placebo got sick more than the ones Drug Users who took the Covid91. We can see that the pink bar for placebo is almost doubled compared to the drug. This indicates the drug is effective for the Drug Users group.

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```
This table shows that the Drug Users who got sick, 33% of them had gotten the drug versus 67% of the sick Drug Users that got the placebo. Based off this information, we can see that the drug seems to be effective for Drug Users. 

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUserMod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is small. This indicates that we need to reject the null hypothesis and that there is a difference between the Drug Users who got the placebo and Drug Users who got the drug. In this case, we can see that the drug was more effective than the placebo, as more Drug Users who got the placebo got sick compared to those that got the drug.


## Calculating Effectiveness

```{r}
DrugUserEstimate <- DrugUserMod$estimate
DrugUserLow <- DrugUserMod$conf.int[1]
DrugUserHigh <- DrugUserMod$conf.int[2]
DrugUserProb <- DrugUserEstimate / (1+DrugUserEstimate)
DrugUserLowProb <- DrugUserLow / (1+DrugUserLow)
DrugUserLowEffect <- DrugUserLowProb * 100
DrugUserHighProb <- DrugUserHigh / (1+DrugUserHigh)
DrugUserHighEffect <- DrugUserHighProb * 100
DrugUserEffect <- DrugUserProb * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Drug Users is `r DrugUserEffect`. The confidence interval for Drug Users is `r DrugUserLowEffect` to `r DrugUserHighEffect`.

# LGBTQ

```{r}
LGBTQ <- filter(Covid91, LGBTQ == "gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According to the bar chart, the LGBTQ community does not benefit from taking the drug. In our sick group, more than 50% had taken the drug. 

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

This table tells us that the LGBTQ should not take this drug. 71% of those who were sick had taken the drug versus the 29% of the sick people who had not taken the drug.

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQMod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small (6e-06). This indicates that we need ti reject the null hypothesis and there is a difference between the LGBTQ who took the placebo and those who took the drug. In this case, the difference is in an ineffective way, as there are more people who got infected while on the drug.

## Calculating Effectiveness

```{r}
LGBTQEstimate <- LGBTQMod$estimate
LGBTQLow <- LGBTQMod$conf.int[1]
LGBTQHigh <- LGBTQMod$conf.int[2]
LGBTQProb <- LGBTQEstimate / (1+LGBTQEstimate)
LGBTQLowProb <- LGBTQLow / (1+LGBTQLow)
LGBTQLowEffect <- LGBTQLowProb * 100
LGBTQHighProb <- LGBTQHigh / (1+LGBTQHigh)
LGBTQHighEffect <- LGBTQHighProb * 100
LGBTQEffect <- LGBTQProb * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for LGBTQ community is `r LGBTQEffect`. The confidence interval for LGBTQ community is `r LGBTQLowEffect` to `r LGBTQHighEffect`.

# Conclusions and recommendations

In conclusion, based off the data we have here, we can see that the effectiveness really does vary among the 4 different subgroups. The bar chart for the Males shows that the Males who took the placebo got sick more than the ones Males who took the Covid91. This indicates the drug is effective for the Males group. According to the barchart, this vaccine is not as effective for Females. The blue bar in the Females chart indicates that out of the Females who are sick more than 50% had gotten the drug. For the Drug Users, the bar chart shows that the Drug Users who took the placebo got sick more than the ones Drug Users who took the Covid91. This indicates the drug is effective for the Drug Users group. As for the LGBTQ community, we can see that the drug is not effective because more than 50% of the sick group got the drug. Reccomendations could be made that Males and Drug Users should take the vaccine, whereas based off the data we have, Females and the LGBTQ community should not get the vaccine, as it seems to not be as effective. 